---
title: "Tsai Lab RNA-seq Pipeline - Part 2"
author: "Poplar Genomics Analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center'
)

# Load pipeline
source("R/config.R")
source("R/utils.R")
source("R/fpkm.R")
source("R/qc.R")
source("R/de.R")
source("R/go.R")

# Initialize
config <- init_pipeline("config/config.yaml")
```

## Overview

This report documents the RNA-seq analysis pipeline including:

- **FPKM Calculation**: Gene expression quantification
- **Sample QC**: PCA and clustering for quality assessment  
- **Differential Expression**: DESeq2 analysis
- **GO Enrichment**: Functional enrichment testing

---

## 1. Data Loading

```{r load-data}
# Load all input data
counts <- load_counts(config)
design <- load_design(config)
comparisons <- load_comparisons(config)

# Display summary
cat("Samples:", nrow(design), "\n")
cat("Genes:", nrow(counts), "\n")
cat("Comparisons:", nrow(comparisons), "\n")
```

### Sample Information

```{r sample-table}
knitr::kable(design, row.names = FALSE)
```

### Comparisons

```{r comparison-table}
knitr::kable(comparisons, row.names = FALSE)
```

---

## 2. FPKM Calculation

```{r fpkm}
fpkm_data <- calculate_fpkm(config)
expressed_genes <- count_expressed_genes(config, fpkm_data)
```

### Expressed Gene Counts

```{r expressed-table}
knitr::kable(expressed_genes, row.names = FALSE, 
             caption = paste0("Genes with FPKM >= ", config$params$fpkm_cutoff))
```

---

## 3. Sample Quality Control

```{r qc, fig.height=10}
qc_results <- run_sample_qc(config, fpkm_data)
```

### PCA Plot

```{r pca-plot}
if (require("ggplot2")) {
  p <- plot_pca(qc_results$pca)
  print(p)
}
```

### Sample Clustering

```{r cluster-plot}
plot(qc_results$clustering, main = "Sample Hierarchical Clustering")
```

---

## 4. Differential Expression

```{r de}
de_results <- run_de_analysis(config, counts, fpkm_data)
```

### DE Summary

```{r de-summary}
knitr::kable(de_results$summary, row.names = FALSE,
             caption = "Differentially Expressed Genes")
```

---

## 5. GO Enrichment

```{r go, eval=config$params$go$enabled}
go_results <- run_go_enrichment(config)
```

GO enrichment analysis completed. Results saved to:
- `data/GO/`: Individual comparison results
- `GO_enrichment_BP_merge.tab`: Biological Process merged results
- `GO_enrichment_MF_merge.tab`: Molecular Function merged results  
- `GO_enrichment_CC_merge.tab`: Cellular Component merged results

---

## Session Info

```{r session}
sessionInfo()
```